---
title: "Getting Started with dismapr"
editor_options: 
  chunk_output_type: console
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE)
```

Example script demonstrating how to use the dismapr package for retrieving
interpolated distance-weighted biomass data from NOAA's DisMAP.

```{r setup}
library(dismapr) 
# devtools::load_all() 
# devtools::install_local(force = T)
library(dplyr)
library(DT)
library(ggplot2)
library(here)
library(mapview)
library(sf)
library(terra)
```

## 1. Indicators

Example 1: Get indicators data for American Lobster in Northeast US Spring

```{r}
tbl_lobster_indicators <- get_dm_indicators(
  common_name = "American lobster", 
  region      = "Northeast US",
  season      = "Spring")
tbl_lobster_indicators

# plot center of gravity changes over time
ggplot(data = tbl_lobster_indicators) +
  geom_point(aes(
    x     = center_of_gravity_longitude, 
    y     = center_of_gravity_latitude, 
    color = year)) +
  labs(
    x     = "Longitude", 
    y     = "Latitude", 
    title = "American Lobster Center of Gravity") +
  theme_classic() +
  theme(
    panel.border = element_rect(
      colour = "black", fill = NA, linewidth = 1))
```

## 2. BioMass Rasters

Example 2: Download interpolated biomass rasters for Summer Flounder in NEUS Spring

```{r}
# show table of dataset_codes for available regions and seasons
dm_regions |> 
  st_drop_geometry() |>
  datatable()

# show map of dataset_codes
# mapView(dm_regions, zcol = "dataset_code") # too big for html
ggplot() +
  geom_sf(
    data = dm_regions, 
    aes(color = dataset_code)) +
  theme_minimal() +
  labs(title = "NOAA DisMAP dataset_code (regions and seasons)") +
  theme(
    panel.border = element_rect(
      colour = "black", fill = NA, linewidth = 1))
```

```{r}
# pick the Northeast US Spring dataset
dataset_code <- "NEUS_SPR"

# show list of available region_seasons (where season may be optional)
get_dm_dataset_layers(dataset_code)

# pick the layer (i.e., species): summer flounder (Paralichthys dentatus)
layer     <- "Paralichthys dentatus"
sp_common <- "SummerFlounder"

# show list of available years
get_dm_dataset_layer_years(dataset_code, layer)

# pick a set of years
years <- 2015:2019

# create directory for downloads
output_dir <- here("inst/test")
dir.create(output_dir, showWarnings = F, recursive = T)

# download each raster
for (yr in years) { # yr = years[1]
  tif <- glue("{output_dir}/{dataset_code}_IDW_{sp_common}_{yr}.tif")

  r <- get_dm_raster(
    dataset_code,
    layer,
    yr,
    tif)
}

yr <- 2015
r <- rast(glue("{output_dir}/{dataset_code}_IDW_{sp_common}_{yr}.tif"))
r
```

### plot interactive, quick

```{r}
plet(r)
```

### plot interactive, with ocean tiles and legend title

```{r}
plet(
  r, 
  tiles = "Esri.OceanBasemap",
  main  = glue(
    "{dataset_code}
    {sp_common}
    {yr}: {names(r)}"))
```

### plot static

```{r}
ggplot() +
  geom_tile(
    data = as.data.frame(r, xy = T) , 
    aes(x = x, y = y, fill = wtcpue)) +
  scale_fill_viridis_c() +
  coord_fixed() +
  labs(title = glue("{dataset_code} {sp_common}, {yr}"))
```

## 3. Survey Locations

```{r}
dataset_code <- "NEUS_SPR"
year <- 2015

pts_yr <- get_dm_survey_locations(dataset_code, where = glue("Year = {year}"))
# show a sample of first 100 rows
head(pts_yr, 100) |> 
  datatable()

# summarize by biomass
pts_yr_sum <- pts_yr |>
  group_by(Longitude, Latitude) |>
  summarize(
    wtcpue = sum(WTCPUE), .groups = "drop")

# plot, static
ggplot() +
  geom_sf(data = pts_yr_sum, aes(color = wtcpue)) +
  theme_minimal() +
  labs(title = glue("Northeast US Spring Survey Locations, {year}"))

# plot, dynamic
mapView(
  pts_yr_sum, zcol = "wtcpue", 
  layer.name = glue("Northeast US Spring Survey Locations, {year}"),
  map.types = c("Esri.OceanBasemap", "OpenStreetMap"))
```

TODO:

- [ ] animated GIF
- [ ] Use [jsonld](https://docs.ropensci.org/jsonld/) to extract from InPort, eg:
  - [Survey Catch-per-unit-effort CURRENT | InPort](https://www.fisheries.noaa.gov/inport/item/69743)
- [ ] explore [`arcgislayers::arc_raster()`](https://developers.arcgis.com/r-bridge/api-reference/arcgislayers/arc_raster.html) for `get_dm_raster()`
- [ ] consider `get_dm_raster_series()` to combine layers years and/or species
